Grappa: Scalable Computing on Mass-Market Hardware for Irregular, Low-Locality Applications

### 2012-07-21 doc/asplos13/relatedwork.tex
related work notes

\subsection{Multithreading to tolerate latency}
hardware: tera, cray
smt,
alewife
gpu,
software threads

@@

Distributed graph processing frameworks like Pregel \cite{pregel:2010} and Distributed GraphLab \cite{distgraphlab:vldb12} provide graph-parallel, vertex-centric programming abstractions that free the application writer from solving distributed system issues like scheduling parallelism, handling fault tolerance, and scaling communication. Pregel adopts a bulk-synchronous parallel (BSP) execution model, which makes it inefficient on workloads that could prioritize vertices. GraphLab, on the other hand, schedules vertex computations individually, allowing prioritization, which gives faster convergence in a variety of iterative algorithms. Grappa also supports dynamic parallelism with asynchronous execution, but has a more general purpose programming model, where parallelism is expressed as tasks or loop iterations. \todo{and express locality? mention pgas e.g. Chapel language implementation? How much to talk about how graphlab achieves scalable performance?}

>>>>

\subsection{Multithreading to tolerate latency}
Using multithreading to tolerate memory latency is well-covered in the literature. Hardware implementations include the Tera MTA \cite{Tera}, Cray XMT \cite{}, Simultaneous multithreading \cite{}, MIT Alewife \cite{}, Cyclops \cite{}, and even GPUs \cite{fatahalian}.

\subsection{systems?}
RAMCloud - low latency; RPC call communication abstraction

sDSM - most are page based and deal with maintaining consistency of data; built on shared memory programming models rather than something more PGAS. They can use prefetching, based on prediction to overlap communication and computation, but does this work for irregular apps? Difference from txDSM is that we don't try to provide txns over multiple nodes, only within one core's memory.

software threads

@@

Distributed graph processing frameworks like Pregel \cite{pregel:2010} and Distributed GraphLab \cite{distgraphlab:vldb12} provide graph-parallel, vertex-centric programming abstractions that free the application writer from solving distributed system issues like scheduling parallelism, handling fault tolerance, and scaling communication. Pregel adopts a bulk-synchronous parallel (BSP) execution model, which makes it inefficient on workloads that could prioritize vertices. GraphLab, on the other hand, schedules vertex computations individually, allowing prioritization, which gives faster convergence in a variety of iterative algorithms. Grappa also supports dynamic parallelism with asynchronous execution, but has a more general purpose programming model, where parallelism is expressed as tasks or loop iterations. \TODO{and express locality? How much to talk about how graphlab achieves scalable performance?}
